\chapter{Nomenclature}
\label{nomenclature}
\begin{adjustbox}{width=.95\textwidth,center}
\noindent\fbox{%
\hfill\parbox{\dimexpr\textwidth-15pt}{%
\vspace{-\topskip}\small
\renewcommand{\entry}[3][\>]{#2 #1 \parbox[t]{.4\textwidth}{\RaggedRight
#3\strut\par}\\}%
\begin{multicols}{2}
  \begin{tabbing}
    \textbf{\large Nomenclature}\\[2ex]
    \textbf{List of variables:}\\
    \entry[\quad\=]{$V_{b}$}{Bulk Volume.}
    \entry{$N_{k}$}{Number of moles for chemical component $k$.}
    \entry{$t$}{Time.}
    \entry{$n_{p}$}{Number of phases present in gridblock.}
    \entry{$x_{kj}$}{Mole fraction of component $k$ in phase $j$}
    \entry{$\xi_{j}$}{Mole density of phase $j$.}
    \entry{$k_{rj}$}{Relative permeability of phase $j$.}
    \entry{$\mu_{j}$}{Viscosity of phase $j$.}
    \entry{$\tensor{K}$}{Permeability tensor.}
    \entry{$\Phi_{j}$}{Gravitational potential of phase $j$.}
    \entry{$\Phi_{w,i}$}{Gravitational potential of water in gridblock $i$.}
    \entry{$\phi$}{porosity.}
    \entry{$S_{j}$}{Saturation of phase $j$.}
    \entry{$\tensor{\Lambda}_{kj}$}{Dispersion tensor of component $k$ in phase $j$.}
    \entry{$q_{k}$}{source or sink term for chemical component $k$.}
    \entry{$\vec{A}_{s}$}{Surface area vector.}
    \entry{$n_{f}$}{Number of neighboring faces.}
    \entry{$\Phi_{i,j}$}{Gravitational potential of face ceneter between gridblock $i$ and $j$.}
    \entry{$\vec{c}_{i,j}$}{Vector from cell $i$ center to face center between gridblock $i$ and $j$.}
    \entry{$\vec{n}_{i,j}$}{Unit vector perpendicular to face area pointing from gridblock $i$ to gridblock $j$.}
    \entry{$T_{i,j}$}{Half-transmissibility between gridblock $i$ and $j$.}
    \entry{$T_{ij}$}{Transmissibility between gridblock $i$ and $j$.}
    \entry{$A_{i,j}$}{Face area between gridblock $i$ and $j$.}
    \entry{$\lambda_{w,ij}$}{Upwind mobility of water for gridblock $i$ and $j$.}
    \entry{$K_{xx},\ K_{xy},\ ...,\ K_{zz}$}{\hspace{2.5cm} Permeability tensor components in a cartesian system.}
    \entry{$J_{F}(x_{n})$}{\hspace{0.5cm}Jacobian matrix of function $F$ evaluated at point $x_{n}$.}
    \entry{$MB_{\alpha}$}{\hspace{0.2cm}Material balance error for phase $\alpha$.}
    \entry{$\bar{B}_{\alpha}$}{Formation-volume factor of phase $\alpha$ evaluated at average pressure.}
    \entry{$\mathcal{R}_{\alpha,i}$}{Residual for phase $\alpha$ evaluated at cell $i$.}
    \entry{$A_{\alpha,i}$}{Accumulation term for phase $\alpha$ evaluated at cell $i$.}
    \entry{$u_{\alpha,ij}$}{Flux for phase $\alpha$ from gridblock $i$ to gridblock $j$.}
    \entry{$f_{\alpha,i}$}{Fugacity of component $i$ in phase $\alpha$.}
    \entry{$N_{p}$}{Total number of phases.}
    \entry{$\rho_{p}$}{Density of phase $p$.}
    \textbf{CPR Specific:}\\
	  \entry{$P^{-1}$}{General two-stage preconditioner.}
	  \entry{$A$}{General matrix to be preconditioned.}
	  \entry{$C$}{A restriction operator used to extract the pressure subsystem \ref{cop}.}
	  \entry{$W^{T}$}{$= C^{T}\cdot DIAG^{-1}(A)$.}
	  \entry{$A_{lm}$}{Sublock of derivatives of $l$ equations with respect to $m$ variable.}
	  \entry{$M_{CPR}^{-1}$}{\hspace{0.5cm}The CPR preconditioning operator.}
	  \entry{$M^{-1}$}{The global preconditioner applied on combined system (usually \texttt{ILU(0)}).}
  \end{tabbing}
\end{multicols}
}%
\hfill}}
\end{adjustbox}
